Modeling the near-wake of a vertical-axis cross-flow turbine with 2-D and 3-D RANS

### 2016-10-02 reviews/reviewer2-response.md

# Response to reviewer 2

We thank the reviewer for his or her very thorough and thoughtful comments,
which have helped us improve the paper.

@@

## Responses to PDF comments

>Effect of this [slip boundary condition] on the results?

It didn't seem to matter much. The mean velocity fields appear to be nearly
symmetrical about the $x$--$y$ plane, but nonetheless, the slip BC is at least
closer to how a free surface would behave than a no slip BC.

>Any studies proving this - PIV works?

To our knowledge, there aren't any studies directly showing that if CFD results
match well at one point in the domain, they will therefore match elsewhere in
the domain, but if the reviewer knows of any, we'd be happy to reference them.

### 2016-10-09 paper.tex
Discuss strut drag modeling effects

The turbine's CAD geometry was prepared or ``cleaned'' for CFD by removing
details determined to be unnecessary, e.g., screw heads and axial shaft grooves,
which would complicate the meshing process and contribute very little to the
overall loading or flow modulation. A drawing of the physically and numerically
modeled geometries is presented in Figure~\ref{fig:cfd-cad}.

@@

@@

Meshes were generated using OpenFOAM's blockMesh and snappyHexMesh utilities.
Mesh  consists of a background hexahedral mesh, which is refined in all
three directions by a factor of 2 in a rectangular region containing the turbine
and near-wake (0.9 m upstream, 1.3 m downstream, $\pm 0.9$ m cross-stream, and
$\pm 0.8$ m vertically). Cells adjacent to the turbine shaft and struts are
refined by a factor of 4, while cells adjacent to the blades are refined by a
factor of 6. To capture the boundary layer, 20 layers were added next to the
blades with an expansion ratio of 1.2. Overall mesh refinement is controlled by
a single parameter---the number of cells in the streamwise direction, $N_x$.

@@

Time step dependence was evaluated using the 2-D $N_x=70$ grid, the results from
which are shown in Figure~\ref{fig:2d-br-verification}. It was seen that the
Spalart--Allmaras model converged well with decreasing time step, leading to a
final time step of 0.001 s. The results from the SST model show a local minimum
at $\Delta t = 0.002$ s, with some divergence for smaller time steps. The local
minimum was chosen as the final time step to run the simulations. Note that the
SST model's convergence behavior may be due to its specific implementation in
OpenFOAM, and not indicative of the nature of the model equations. Verification
studies for CFTs with this level of detail in the literature are not common,
though the final time step is comparable to others \cite{Balduzzi2016}. Note
that while in general there may be a coupling between necessary spatial and
temporal resolution, these values were arrived at iteratively, i.e., the spatial
dependence was varied with the chosen time step, and vice versa, such that there
was one independent variable for each parameter sweep.

@@

Turbine operation in all cases was simulated for 10 seconds, or approximately
six rotor revolutions. Computations for the 3-D cases were run on 192 processes
(24 nodes $\times$ 8 cores each on Sandia National Labs' Red Mesa high
performance computing cluster) and took on the order of 1,000 CPU hours per
second of simulated time. The 2-D simulations were run on a single processor and
took on the order of one CPU hour per simulated second.

@@

The 3-D simulations matched the experimental measurements more closely. The
Spalart--Allmaras model's mean power coefficient was 0.24 while the SST's was
0.33, which represent a 6\% under- and 30\% overprediction, respectively,
compared with the experiments. Some of the apparent overprediction of rotor drag
coefficient could also be an effect of the experimental procedure, where the
``tare drag'' from the turbine mounting structure was measured without a turbine
installed, then subtracted in post-processing. With the turbine installed, flow
past the mounting frame will be higher due to blockage, meaning the tare drag
would be underestimated, and the CFD results for $C_D$ should be slightly higher
than the experimental measurements.

@@

Turbulence kinetic energy contours for the experimental measurements and each
CFD case at $x/D=1$ are presented in Figure~\ref{fig:br-cfd-kcont}. As seen in
the profiles in Figure~\ref{fig:br-cfd-profiles}, the SA model was not able to
resolve the majority of the unsteadiness in the flow. In contrast, the SST model
did a good job predicting the locations and magnitudes of various peaks in $k$.
These are generated along the top of the turbine via tip vortex shedding, and
the $-y$ side of the turbine via dynamic stall. We do however see the smearing
effect from the dynamic stall vortex centered around $z/H=0$, which is likely
more of an issue with wake evolution rather than wake generation.

>>>>

The turbine's CAD geometry was prepared or ``cleaned'' for CFD by removing
details determined to be unnecessary, e.g., screw heads and axial shaft grooves,
which would complicate the meshing process and contribute very little to the
overall loading or flow modulation. A drawing of the physically and numerically
modeled geometries is presented in Figure~\ref{fig:cfd-cad}, and geometric
parameters are listed in Table~\ref{tab:turb-geom}.

@@

\begin{table}
    \centering
    \begin{tabular}{r|l}
        Parameter & Value \\
        \hline
        Diameter $D$ & 1.0 m \\
        Height or blade span $H$ & 1.0 m \\
        Blade chord length $c$ & 0.14 m \\
        Blade profile & NACA 0020 \\
        Blade mounting point & $c/2$ and $H/2$ \\
        Strut chord length & 0.14 m \\
        Strut profile & NACA 0020
    \end{tabular}

    \caption{Geometric parameters for the UNH-RVAT rotor modeled in this study.}

    \label{tab:turb-geom}
\end{table}

@@

Meshes were generated using OpenFOAM's blockMesh and snappyHexMesh utilities.
The mesh consisted of a background hexahedral mesh, which was refined in all
three directions by a factor of 2 in a rectangular region containing the turbine
and near-wake (0.9 m upstream, 1.3 m downstream, $\pm 0.9$ m cross-stream, and
$\pm 0.8$ m vertically). Cells adjacent to the turbine shaft and struts were
refined by a factor of 4, while cells adjacent to the blades were refined by a
factor of 6. To capture the boundary layer, i.e., where flow physics most affect
rotor torque---20 layers were added next to the blades with an expansion ratio
of 1.2. Overall mesh refinement is controlled by a single parameter---the number
of cells in the streamwise direction, $N_x$.

@@

Time step dependence was evaluated using the 2-D $N_x=70$ grid, the results from
which are shown in Figure~\ref{fig:2d-br-verification}. It was seen that the
Spalart--Allmaras model converged well with decreasing time step, leading to a
final time step of 0.001 s. The results from the SST model show a local minimum
at $\Delta t = 0.002$ s, with some divergence for smaller time steps, the reason
for which is unknown. The local minimum was chosen as the final time step to run
the simulations. Note that the SST model's convergence behavior may be due to
its specific implementation in OpenFOAM, and not indicative of the nature of the
model equations. Note that while in general there may be a coupling between
necessary spatial and temporal resolution, these values were arrived at
iteratively, i.e., the spatial dependence was varied with the chosen time step,
and vice versa, such that there was one independent variable for each parameter
sweep. Reassuringly, the chosen time step and average $y^+$ adjacent to the
blades are comparable to others reported in the literature
\cite{Maitre2013,Balduzzi2016}.

@@

Turbine operation in all cases was simulated for 10 seconds, or approximately
six rotor revolutions, to allow the revolution-averaged performance coefficient
to converge to within $\sim 1$ percentage point; cf.
Figure~\ref{fig:cp-convergence}. Computations for the 3-D cases were run on 192
processes (24 nodes $\times$ 8 cores each on Sandia National Labs' Red Mesa high
performance computing cluster) and took on the order of 1,000 CPU hours per
second of simulated time. The 2-D simulations were run on a single processor and
took on the order of one CPU hour per simulated second.

    \includegraphics[width=0.75\textwidth]{cp-time-series}

    \caption{Revolution-averaged power coefficient versus rotor revolution
    simulated with the 3-D $k$--$\omega$ SST model.}

    \label{fig:cp-convergence}
\end{figure}

@@

The 3-D simulations matched the experimental measurements more closely. The
Spalart--Allmaras model's mean power coefficient was 0.24 while the SST's was
0.33, which represent a 6\% under- and 30\% overprediction, respectively,
compared with the experiments. Strut drag modeling discrepancies due to the
struts' lower mesh resolution compared to the blades, along with the use of wall
functions, could have played a minor role in degrading overall performance
predictions, and should be investigated further.

Some of the apparent overprediction of rotor drag coefficient could also be an
effect of the experimental procedure, where the ``tare drag'' from the turbine
mounting structure was measured without a turbine installed, then subtracted in
post-processing. With the turbine installed, flow past the mounting frame will
be higher due to blockage, meaning the tare drag would be underestimated, and
the CFD results for $C_D$ should be slightly higher than the experimental
measurements.

@@

Turbulence kinetic energy contours for the experimental measurements and each
CFD case at $x/D=1$ are presented in Figure~\ref{fig:br-cfd-kcont}. As seen in
the profiles in Figure~\ref{fig:br-cfd-profiles}, the SA model was not able to
resolve the majority of the unsteadiness in the flow. In contrast, the SST model
did a good job predicting the locations and magnitudes of various peaks in $k$.
These are generated along the top of the turbine via tip vortex shedding, and
the $-y$ side of the turbine via dynamic stall. We do however see the smearing
effect from the dynamic stall vortex centered around $z/H=0$, which is likely
more of an issue with wake evolution rather than wake generation. The effects of
mesh resolution on the breakdown and diffusion of the dynamic stall vortex may
also have an influence here.
